---
title: "TIME of PTC and ATC"
author: "fengchao"
date: "`r Sys.Date()`"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, eval = F,root.dir = '~/Desktop/thyroid/thyroid-srna-seq/')
```

## Some preparation

### Set color

```{r}
condition_col = c('#009245','#F9AE7D','#BE1E2D')
```

### Load package

```{r}
library(Seurat)
library(ggpubr)
```

### Load data

#### Load NBT

```{r}
source('./code/utils.R')
files = list.files('./raw/scRNA-seq/NBT/')
files = files[grep('.*ATC.*',files)]

for (file in files) {
  ID = str_extract(file,'ATC\\d+')
  print(ID)
  tmp = load_NBT_data(str_c('./raw/scRNA-seq/NBT/',file),ID)
  assign(ID,tmp)
  
}

```

#### Load NC

```{r}
clinical <- read_tsv('./raw/scRNA-seq/NC/meta.tsv',col_names = F)
colnames(clinical) = c('GSM','type')
PTC_clinical = clinical %>% 
  filter(str_detect(type,' tumor'))
rm(clinical)
print(PTC_clinical$GSM)
PTC_clinical$myID <- str_c('PTC',1:7)

dirs = str_c('./raw/scRNA-seq/NC/',PTC_clinical$GSM)

for (i in 1:length(dirs)) {
  file = dirs[i]
  gsm_num = PTC_clinical$myID[i]
  print(file)
  seurat_object <- CreateSeuratObject(Read10X(data.dir = file),
                                      project = gsm_num)
  
  seurat_object[["percent.mt"]] <- PercentageFeatureSet(seurat_object,
                                                        pattern = "^MT-")
  
  seurat_object[["percent.rp"]] <- PercentageFeatureSet(seurat_object,
                                                        pattern = "^RP[SL]")
  seurat_object <- subset(seurat_object, 
                          subset = nFeature_RNA > 300 & 
                            nFeature_RNA < 2000 & 
                            percent.mt < 10 & percent.rp < 40)
  assign(gsm_num,seurat_object)
}

rm(seurat_object,file,gsm_num)
```

### All cells

#### Merge

```{r}
se_ob <- merge(x =PTC1,
               y = c(PTC2,PTC3,PTC4,PTC5,PTC6,PTC7,ATC1,ATC2,ATC3,ATC4,ATC5),
               add.cell.ids = c(str_c('PTC',1:7),
                                str_c('ATC',1:5)),
               project = "PTC.ATC")
saveRDS(se_ob,file = '')
```

#### Workflow

```{r}
# Workflow ----
se_ob <- Seurat::NormalizeData(se_ob, 
                               normalization.method = "LogNormalize", 
                               scale.factor = 10000) %>% 
  FindVariableFeatures(., 
                       selection.method = "vst", 
                       nfeatures = 2000)

# all.genes <- rownames(se_ob)
se_ob <- ScaleData(se_ob)

se_ob <- RunPCA(se_ob,
                   features = VariableFeatures(object = se_ob))

DimPlot(se_ob, reduction = "pca")
ElbowPlot(se_ob)
VlnPlot(object = se_ob, features = "PC_1", group.by = "orig.ident", pt.size = .1)

library(harmony)
# Remove batch effect ----
se_ob <- se_ob %>% 
  RunHarmony("orig.ident", plot_convergence = TRUE)

library(cowplot)
p1 = DimPlot(se_ob, reduction = "harmony")
p2 = ElbowPlot(se_ob,reduction = "harmony",ndims = 30)
p3 = VlnPlot(object = se_ob, features = "harmony_1", group.by = "orig.ident", pt.size = .1)
plot_grid(p1,p2,p3)

se_ob <- RunUMAP(se_ob, 
                 reduction = 'harmony', 
                 dims = 1:20,
                 seed.use = 1) %>% 
  RunTSNE(.,reduction = 'harmony', 
          dims = 1:20,
          seed.use = 1) %>% 
  FindNeighbors(., reduction = 'harmony', 
                dims = 1:20)

se_ob <- FindClusters(se_ob, 
                      resolution = 0.8) 
```

#### Del Doublet

```{r}
source('~/Documents/mytools/seurat_help_func.R')
se_ob_list <- SplitObject(se_ob, split.by = "orig.ident")
se_ob_list <- lapply(se_ob_list,seurat_workflow)

se_ob <- scCustomize::Merge_Seurat_List(se_ob_list,
                                        project = 'PTC.ATC')
se_ob <- subset(se_ob,DF.classifications != 'Doublet')
```

#### QC and Ribosomal gene

```{r}

se_ob_list <- SplitObject(se_ob, split.by = "orig.ident")
se_ob <- scCustomize::Merge_Seurat_List(se_ob_list,
                                        project = 'PTC.ATC')
rm(se_ob_list)
gc()

se_ob@meta.data$condition = str_sub(se_ob$orig.ident,1,3)
se_ob[["percent.mt"]] <- PercentageFeatureSet(se_ob,
                                              pattern = "^MT-")

se_ob[["percent.rp"]] <- PercentageFeatureSet(se_ob,
                                              pattern = "^RP[SL]")

VlnPlot(se_ob, 
        features = c("nFeature_RNA", "nCount_RNA", "percent.mt"), 
        ncol = 4,
        pt.size = 0.,
        group.by = 'orig.ident')

# Filter out low quality cells ----
se_ob <- subset(se_ob, 
                subset = nFeature_RNA > 300 & nFeature_RNA < 7000 & percent.mt < 10)
```

#### Workflow after Del doublet

```{r}
# Workflow ----
se_ob <- Seurat::NormalizeData(se_ob, 
                               normalization.method = "LogNormalize", 
                               scale.factor = 10000) %>% 
  FindVariableFeatures(., 
                       selection.method = "vst", 
                       nfeatures = 2000)

# all.genes <- rownames(se_ob)
se_ob <- ScaleData(se_ob)

se_ob <- RunPCA(se_ob,
                   features = VariableFeatures(object = se_ob))

ElbowPlot(se_ob,ndims = 50)

library(harmony)
# Remove batch effect ----
se_ob <- se_ob %>% 
  RunHarmony("orig.ident", plot_convergence = TRUE)

library(cowplot)
p1 = DimPlot(se_ob, reduction = "harmony")
p2 = ElbowPlot(se_ob,reduction = "harmony",ndims = 40)
p3 = VlnPlot(object = se_ob, features = "harmony_1", group.by = "orig.ident", pt.size = .1)
plot_grid(p1,p2,p3)

se_ob <- RunUMAP(se_ob, 
                 reduction = 'harmony', 
                 dims = 1:20,
                 seed.use = 1) %>% 
  RunTSNE(.,reduction = 'harmony', 
          dims = 1:20,
          seed.use = 1) %>% 
  FindNeighbors(., reduction = 'harmony', 
                dims = 1:20)

se_ob <- FindClusters(se_ob, 
                      resolution = c(0.2, 0.4, 0.6,
                                     0.8, 1.0, 1.2)) 
```

#### Dimplot

```{r}
DimPlot(se_ob,group.by = 'RNA_snn_res.1',label = T) 
```

#### Marker Gene

```{r}
# immune cells 
imm_gene = c("PTPRC",'CD3D',"CD79A")
# stromal cells 
fibo = c('MME','ACTA2')
en = c('PECAM1')
# epithelial cells 
ep = c('EPCAM')
thyroid = c('TG','TSHR','TPO','KRT19','FN1','CITED1')
genes_to_check = c(imm_gene,ep,thyroid,fibo,en)

library(scRNAtoolVis)

jjDotPlot(object = se_ob,
          gene = genes_to_check,
          ytree = T,
          rescale = T,
          rescale.min = 0, 
          rescale.max = 1,
          id = 'RNA_snn_res.0.6')
```

#### Annotation

```{r}
imm = c(2,3,4,5,6,7,8,9,16,17,19,22)
ep = c(0,10,12,13,15)
stro = c(1,11,14,18,20,21)
cell_map <- tibble(seurat_clusters = c(imm,ep,stro),
                   main_type = c(rep('immune',length(imm)),
                                 rep('epithelial',length(ep)),
                                 rep('stromal',length(stro))))
cell_map <- cell_map %>% 
  arrange(seurat_clusters)

cluster_letters <- cell_map$main_type[Idents(object = se_ob)]
se_ob <- AddMetaData(
  object = se_ob,
  metadata = cluster_letters,
  col.name = 'cell_types_1')
DimPlot(se_ob,group.by = 'cell_types_1')
```

#### Del Some Clusters

> 10,23,25,26 looks like doublet

```{r}
se_ob <- subset(se_ob,RNA_snn_res.0.6 %in% c(0,1,2,3,4,5,6,7,8,9,
                                             11,12,14,15,16,17,18,19,20,
                                             21,22))
# re-run the workflow
source('~/Documents/mytools/seurat_help_func.R')
```

#### Re-order Clusters

```{r}
se_ob$re_order_cluster_0.6 <- factor(se_ob$RNA_snn_res.0.6,
                                     levels = c(2,3,4,5,6,7,8,9,16,17,19,22,
                                                0,10,12,13,15,
                                                1,11,14,18,20,21))
cell_types <- c(str_c('immune-C',1:length(c(2,3,4,5,6,7,8,9,16,17,19,22))),
                str_c('epithelial-C',1:length(c(0,12,15))),
                str_c('stromal-C',1:length(c(1,11,14,18,20,21))))
cluster_letters <- cell_types[factor(as.character(se_ob$re_order_cluster_0.6),
                                     levels = c(2,3,4,5,6,7,8,9,16,17,19,22,
                                                0,12,15,
                                                1,11,14,18,20,21))]
cluster_letters <- factor(cluster_letters,
                          levels = cell_types)
se_ob <- AddMetaData(
  object = se_ob,
  metadata = cluster_letters,
  col.name = 'cell_types_2')
```

#### Fig S1

```{r}
####
col = c("#ffc9d3", "#ffccd4", "#f2aa96", "#ed6f8e", "#f9b6ce", "#ff9d9b",
        "#ffc2bc", "#f28e9f", "#ffd6cc", "#ed7d93", "#f78fb2", "#ffccd7",
        "#c4fbff", "#7ad3f9", "#9feaf4", "#97f285", "#6ded8f", "#b6ff7a",
        "#9ef492", "#b0fc8f", "#b0f79e")
DimPlot(se_ob,
        group.by = 'cell_types_2',
        cols = col,
        label = T,
        label.size = 3)
####
jjDotPlot(object = se_ob,
          gene = genes_to_check,
          ytree = F,
          rescale = T,
          rescale.min = 0, 
          rescale.max = 1,
          id = 'cell_types_2')
####
col = c(randomcoloR::distinctColorPalette(5),
        randomcoloR::distinctColorPalette(7))
DimPlot(se_ob,
        group.by = 'orig.ident',
        cols = col,
        label = F,
        label.size = 3)
####
cell.ident <- data.frame(orig.ident = se_ob$orig.ident,
                        sample = se_ob$condition,
                        cluster = se_ob$seurat_clusters,
                        ann    = se_ob$cell_types_1)

plotData <- cell.ident %>% 
  group_by(orig.ident,ann) %>% 
  summarise(count = n()) %>% 
  group_by(orig.ident) %>% 
  mutate(freq = round(count/(sum(count)),3))
library(ggpubr)
ggbarplot(plotData, "orig.ident", "freq",
          fill = "ann", color = "ann", palette = "aaas",
          label = F, lab.pos = "in")+
  theme(legend.position = 'right')+
  rotate_x_text(90)+xlab('')
```

### Immune

#### Subset Immune Cells

```{r}
se_ob_imm <- subset(se_ob,cell_types_1 == 'immune')
```

#### Workflow

```{r}
# Workflow ----
se_ob_imm <- Seurat::NormalizeData(se_ob_imm, 
                               normalization.method = "LogNormalize", 
                               scale.factor = 10000) %>% 
  FindVariableFeatures(., 
                       selection.method = "vst", 
                       nfeatures = 2000)

# all.genes <- rownames(se_ob_imm)
se_ob_imm <- ScaleData(se_ob_imm)

se_ob_imm <- RunPCA(se_ob_imm,
                   features = VariableFeatures(object = se_ob_imm))

DimPlot(se_ob_imm, reduction = "pca")
ElbowPlot(se_ob_imm,ndims = 50)
VlnPlot(object = se_ob_imm, features = "PC_1", group.by = "orig.ident", pt.size = .1)

library(harmony)
# Remove batch effect ----
se_ob_imm <- se_ob_imm %>% 
  RunHarmony("orig.ident", plot_convergence = TRUE)

library(cowplot)
p1 = DimPlot(se_ob_imm, reduction = "harmony")
p2 = ElbowPlot(se_ob_imm,reduction = "harmony",ndims = 30)
p3 = VlnPlot(object = se_ob_imm, features = "harmony_1", group.by = "orig.ident", pt.size = .1)
plot_grid(p1,p2,p3)

se_ob_imm <- RunUMAP(se_ob_imm, 
                 reduction = 'harmony', 
                 dims = 1:20,
                 seed.use = 1) %>% 
  RunTSNE(.,reduction = 'harmony', 
          dims = 1:20,
          seed.use = 1) %>% 
  FindNeighbors(., reduction = 'harmony', 
                dims = 1:20)

se_ob_imm <- FindClusters(se_ob_imm, 
                      resolution = c(0.2, 0.4, 0.6,
                                     0.8, 1.0, 1.2)) 
```

#### Annotation Marker Gene

```{r}
nk.marker <- c('NKG7','GNLY','KLRD1','KLRF1')
t.marker <- c('CD3D','CD8A','CD4','CD247')
b.marker <- c('CD79A', 'MS4A1', 'IGKC', 'IGHG1', 'MZB1')
macphage.marker <- c('CD68','CD14', 'FCGR3A', 'LYZ', 'S100A8', 'S100A9','CD163')
dc.marker <- c('HLA-DRA', 'HLA-DRB1')

imm.makrer <- c(nk.marker,t.marker,b.marker,macphage.marker,dc.marker)

DimPlot(se_ob_imm,
        group.by = 'RNA_snn_res.1.2',
        label = T,
        cols = randomcoloR::randomColor(24))

jjDotPlot(object = se_ob_imm,
          gene = imm.makrer,
          xtree = F,
          ytree = T, 
          # rescale = T, rescale.min = 0, rescale.max = 1,
          id = 'RNA_snn_res.1.2')
```

#### Annotation Cell Type

```{r}
nk = c(8)
t = c(0,1,3,4,5,6,15,16,17,19,21)
myeloid = c(2,7,9,11,13,20,22)
b = c(10,12,14,18,23) 

cell_map = data.frame(clusters     = c(nk,t,myeloid,b),
                      
                      cell_types_3 = c(rep('NK',length(nk)),
                                       rep('T',length(t)),
                                       rep('Myeloid',length(myeloid)),
                                       rep('B',length(b))),
                      
                      cell_types_4 = c(str_c('NK-C',1:length(nk)),
                                       str_c('T-C',1:length(t)),
                                       str_c('Myeloid-C',1:length(myeloid)),
                                       str_c('B-C',1:length(b))),
                      my_cluster   = c(0,1,2,3,4,5,6,7,8,9,10,
                                       11,12,13,14,15,1,1,18,0,7,
                                       1,2,12))

cell_map <- cell_map %>% 
  arrange(clusters)

cluster_letters <- cell_map$cell_types_3[se_ob_imm$RNA_snn_res.1.2]

cluster_letters <- factor(cluster_letters,
                          levels = c('NK','T','Myeloid','B'))

se_ob_imm <- AddMetaData(
  object = se_ob_imm,
  metadata = cluster_letters,
  col.name = 'cell_types_3')

cluster_letters <- cell_map$cell_types_4[se_ob_imm$RNA_snn_res.1.2]

cluster_letters <- factor(cluster_letters,
                          levels = c(str_c('NK-C',1:length(nk)),
                                     str_c('T-C',1:length(t)),
                                     str_c('Myeloid-C',1:length(myeloid)),
                                     str_c('B-C',1:length(b))))

se_ob_imm <- AddMetaData(
  object = se_ob_imm,
  metadata = cluster_letters,
  col.name = 'cell_types_4')

DimPlot(se_ob_imm,
        group.by = 'cell_types_3')

DimPlot(se_ob_imm,
        group.by = 'cell_types_4',
        cols = randomcoloR::randomColor(24),
        label = T,label.size = 3)
####
col = c(randomcoloR::randomColor(5),
        randomcoloR::randomColor(7))
DimPlot(se_ob_imm,
        group.by = 'orig.ident',
        cols = col,
        label = F,
        label.size = 3)
```

#### Fig

```{r}
####
cell.ident <- data.frame(orig.ident = se_ob_imm$orig.ident,
                        sample = se_ob_imm$condition,
                        cluster = se_ob_imm$seurat_clusters,
                        ann    = se_ob_imm$cell_types_3)

plot.data <- cell.ident %>% 
  group_by(orig.ident,ann) %>% 
  summarise(count = n()) %>% 
  group_by(orig.ident) %>% 
  mutate(freq = round(count/(sum(count)),3))
library(ggpubr)
ggbarplot(plot.data, "orig.ident", "freq",
          fill = "ann", color = "ann", palette = "aaas",
          label = F, lab.pos = "in")+
  theme(legend.position = 'right')+
  rotate_x_text(90)+xlab('')
```

### T & NK cells

#### Subset T & NK cells

```{r}
se_ob_t <- subset(se_ob_imm,cell_types_3 %in% c('NK','T'))
source('~/Documents/mytools/seurat_help_func.R')
```

#### Workflow

```{r}
# Workflow ----
se_ob_t <- Seurat::NormalizeData(se_ob_t, 
                               normalization.method = "LogNormalize", 
                               scale.factor = 10000) %>% 
  FindVariableFeatures(., 
                       selection.method = "vst", 
                       nfeatures = 2000)

# all.genes <- rownames(se_ob_t)
se_ob_t <- ScaleData(se_ob_t)

se_ob_t <- RunPCA(se_ob_t,
                   features = VariableFeatures(object = se_ob_t))

DimPlot(se_ob_t, reduction = "pca")
ElbowPlot(se_ob_t,ndims = 50)
VlnPlot(object = se_ob_t, features = "PC_1", group.by = "orig.ident", pt.size = .1)

library(harmony)
# Remove batch effect ----
se_ob_t <- se_ob_t %>% 
  RunHarmony("orig.ident", plot_convergence = TRUE)

library(cowplot)
p1 = DimPlot(se_ob_t, reduction = "harmony")
p2 = ElbowPlot(se_ob_t,reduction = "harmony",ndims = 30)
p3 = VlnPlot(object = se_ob_t, features = "harmony_1", group.by = "orig.ident", pt.size = .1)
plot_grid(p1,p2,p3)

se_ob_t <- RunUMAP(se_ob_t, 
                 reduction = 'harmony', 
                 dims = 1:20,
                 seed.use = 1) %>% 
  RunTSNE(.,reduction = 'harmony', 
          dims = 1:20,
          seed.use = 1) %>% 
  FindNeighbors(., reduction = 'harmony', 
                dims = 1:20)

se_ob_t <- FindClusters(se_ob_t, 
                      resolution = c(0.2, 0.4, 0.6,
                                     0.8, 1.0, 1.2)) 
```

#### Annotation Marker Gene

```{r}
markers = c('CD4','CD8A',
            'IL2RA','FOXP3','CTLA4',
            'HAVCR2','LAG3','PDCD1','TIGIT',
            'GZMB','GNLY','IFNG','KLRD1',
            'SELL','CCR7','TCF7','LEF1',
            'TRGV5')

jjDotPlot(object = se_ob_t,
          gene = markers,
          xtree = F,
          ytree = T, 
          # rescale = T, rescale.min = 0, rescale.max = 1,
          id = 'RNA_snn_res.1.2')

DimPlot(se_ob_t,
        group.by = 'RNA_snn_res.1.2',
        label = T,
        cols = randomcoloR::randomColor(24))

DimPlot(se_ob_t,
        group.by = 'RNA_snn_res.1.2',
        split.by = 'orig.ident',
        label = T,
        cols = randomcoloR::randomColor(24))
```

#### Annotation Cell Type

```{r}
cd4 <- c(1,2,8,14,15)
cd8 <- c(0,3,5,6,9,10,11,12,13)
nk <- c(4)
tgd <- c(7,16)

cell_types_6 <- c(rep('CD4',length(cd4)),
                 rep('CD8',length(cd8)),
                 rep('NK',length(nk)),
                 rep('Tgd',length(tgd)))
cell_type_5 <- c('CD4-C1','CD4-C2','CD4-C3','CD4-*','CD4-*',
                 'CD8-C1','CD8-C2','CD8-C3','CD8-C4','CD8-C5',
                 'CD8-C6','CD8-C7','CD8-C8','CD8-C9',
                 'NK',
                 'Tgd','Tgd')
cell_map <- data.frame(clusters = c(cd4,cd8,nk,tgd),
                       cell_types_6 = cell_types_6,
                       cell_types_5 = cell_type_5) %>% 
  arrange(clusters)


cluster_letters <- cell_map$cell_types_6[se_ob_t$RNA_snn_res.1.2]
se_ob_t <- AddMetaData(
  object = se_ob_t,
  metadata = cluster_letters,
  col.name = 'cell_types_6')

cluster_letters <- cell_map$cell_types_5[se_ob_t$RNA_snn_res.1.2]
se_ob_t <- AddMetaData(
  object = se_ob_t,
  metadata = cluster_letters,
  col.name = 'cell_types_5')

DimPlot(se_ob_t,
        group.by = 'cell_types_6',
        label = T,label.size = 3.5,
        cols = randomcoloR::randomColor(17))

jjDotPlot(object = se_ob_t,
          gene = markers,
          xtree = F,
          ytree = F, 
          rescale = T, rescale.min = 0, rescale.max = 1,
          id = 'cell_types_5',
          cluster.order = c('CD4-C1','CD4-C3','CD4-C2',
                            'CD8-C1','CD8-C8','CD8-C6','CD8-C3','CD8-C2',
                            'CD8-C5','CD8-C9','CD8-C7','CD8-C4',
                            'Tgd','NK'))
```

#### Del CD4-\* cluster

```{r}
seurat_save_rds(se_ob_t,path = './data/PTC_ATC/se_ob_t.rds')
se_ob_t <- subset(se_ob_t,
                  cell_types_5 %in% c('CD4-C1','CD4-C2','CD4-C3',
                                      'CD8-C1','CD8-C2','CD8-C3','CD8-C4','CD8-C5',
                                      'CD8-C6','CD8-C7','CD8-C8','CD8-C9',
                                      'NK',
                                      'Tgd','Tgd'))
seurat_save_rds(se_ob_t,path = './data/PTC_ATC/se_ob_t_del_CD4-*.rds')


DimPlot(se_ob_t,
        group.by = 'cell_types_5',
        label = T,label.size = 3.5,
        cols = randomcoloR::randomColor(17))
jjDotPlot(object = se_ob_t,
          gene = markers,
          xtree = F,
          ytree = F, 
          rescale = T, rescale.min = 0, rescale.max = 1,
          id = 'cell_types_5',
          cluster.order = c('CD4-C1','CD4-C3','CD4-C2',
                            'CD8-C1','CD8-C8','CD8-C6','CD8-C3','CD8-C2',
                            'CD8-C5','CD8-C9','CD8-C7','CD8-C4',
                            'Tgd','NK'))
```

#### FindMarkers

```{r}
Idents(se_ob_t) <- 'cell_types_5'
cd8.t.markers <- FindAllMarkers(subset(se_ob_t,cell_types_6 != 'CD4'),
                            test.use = 'MAST',
                            only.pos = T)
cd4.t.markers <- FindAllMarkers(subset(se_ob_t,cell_types_6=='CD4'),
                            test.use = 'MAST',
                            only.pos = T)
```

#### Cal cyto exhau score

```{r}
t.genesets <- readxl::read_excel('~/Documents/bio-ann-file/T_genesets_from_nature_medicine/Table.S6.xlsx')
t.genesets <- as.list(t.genesets)
t.genesets <- lapply(t.genesets, function(x){
                                              x <- x[!is.na(x)] 
                                              return(x)})
se_ob_t <- AddModuleScore(
  object = se_ob_t,
  features = list(t.genesets$Cytotoxicity),
  name = 'Cytotoxic'
)
se_ob_t <- AddModuleScore(
  object = se_ob_t,
  features = list(t.genesets$Exhaustion),
  name = 'Exhaution'
)
se_ob_t <- AddModuleScore(
  object = se_ob_t,
  features = list(t.genesets$Naïve),
  name = 'Naive'
)
se_ob_t <- AddModuleScore(
  object = se_ob_t,
  features = list(treg_gene$SYMBOL),
  name = 'Treg'
)
```

#### Cal cell cycle score

```{r}
s.genes <- cc.genes$s.genes
g2m.genes <- cc.genes$g2m.genes

se_ob_t <- CellCycleScoring(se_ob_t, 
                            s.features = s.genes, 
                            g2m.features = g2m.genes, 
                            set.ident = TRUE)
```

#### Cell Fraction

```{r}
cell_ident <- data.frame(orig.ident = se_ob_t$orig.ident,
                        sample = se_ob_t$condition,
                        cluster = se_ob_t$cell_types_5)

plot.data <- cell_ident %>% 
  group_by(orig.ident,sample,cluster) %>% 
  summarise(count = n()) %>% 
  ungroup() %>% 
  group_by(orig.ident) %>% 
  mutate(freq = round(count/(sum(count)),3)) %>% 
  ungroup() 

# bar plot
ggbarplot(plot.data,'cluster','freq',add = "median_q1q3",
          fill = 'sample',palette = cdt_color,
          position = position_dodge(0.9))+
  rotate_x_text(45)
```

#### my ECDF(用Boxplot代替)

```{r}
plot.data <- se_ob_t@meta.data %>% 
  dplyr::select(.,c('condition','cell_types_5',
                    'Cytotoxic1','Exhaution1','Naive1','Treg1')) %>% 
  dplyr::rename(group = cell_types_5,
                cytotoxic = Cytotoxic1,
                exhaution = Exhaution1,
                treg = Treg1,
                naive = Naive1)

plot.data %>% 
  group_by(group) %>% 
  summarise(naive = median(naive)) %>% 
  arrange(naive) %>% 
  ggbarplot(x='group',y='naive',
            fill = 'group',
            palette = randomcoloR::randomColor(14))+
  rotate_x_text(40)

plot.data %>%
  mutate(group2 = str_sub(group,1,3)) %>% 
  ggboxplot(.,
          x= 'condition',
          y = 'cytotoxic',
          title = 'cd8.t',outlier.shape = NA)+
  facet_wrap(facets = 'group2',scales = 'free',nrow = 1)+
  stat_compare_means(comparisons = list(c('PTC','ATC')))
```

#### ssGSVA

```{r}

library(Seurat)
library(msigdbr)
library(GSVA)
library(clusterProfiler)
library(patchwork)
scRNA <- subset(se_ob_t,cell_types_6 == 'CD4')
DefaultAssay(scRNA) <- "RNA"
scRNA <- NormalizeData(scRNA)

Idents(scRNA) <- "cell_types_5" 
expr <- AverageExpression(scRNA, assays = "RNA", slot = "data")[[1]]
expr <- expr[rowSums(expr)>0,]  
expr <- as.matrix(expr)

## geneset1
genesets <- msigdbr(species = "Homo sapiens", category = "H") 
genesets <- subset(genesets, select = c("gs_name","gene_symbol")) %>% 
  as.data.frame() %>% 
  mutate(gs_name = str_remove(gs_name,'HALLMARK_'))
seleted_pathway = c('APOPTOSIS','HYPOXIA','EPITHELIAL_MESENCHYMAL_TRANSITION',
                    'GLYCOLYSIS','ANGIOGENESIS')
genesets <- genesets %>% dplyr::filter(gs_name %in% seleted_pathway)
genesets <- split(genesets$gene_symbol, genesets$gs_name)

## geneset2
genesets2 <- readxl::read_excel('~/Documents/bio-ann-file/2499_Immune_Related_Gene.xlsx') 
# genesets2$Symbol <- str_to_title(genesets2$Symbol)
genesets2 <- split(genesets2$Symbol,genesets2$Category)
# genesets2$BCRSignalingPathway = NULL
# genesets2$TCRsignalingPathway = NULL
# genesets2$NaturalKiller_Cell_Cytotoxicity = NULL

genesets = c(genesets,genesets2)

gsva.res <- GSVA::gsva(expr, genesets, method="gsva") 

p_macro = pheatmap::pheatmap(gsva.res, 
                             show_colnames = T,show_rownames = T,
                             scale = "row",
                             cluster_cols=T,
                             cluster_rows = T,
                             main = 'CD4',
                             # labels_col = c('nThyroid','PTC','ATC')
                             )
```

### CD8 T

```{r}
se_ob_cd8.t <- subset(se_ob_t,cell_types_6 == 'CD8')
```

#### my ECDF

```{r}
source('~/Documents/mytools/my_ecdf.r')
```

#### Fig3 b Dotplot

```{r}
jjDotPlot(object = se_ob_cd8.t,
          gene = c('Exhaution1','Cytotoxic1','Naive1','G2M.Score','ENTPD1'),
          ytree = T,
          rescale = T,
          rescale.min = 0, 
          rescale.max = 1,
          id = 'cell_types_5')
```

#### Fig3 c

```{r}
se_ob_cd8.t@meta.data %>% 
  group_by(cell_types_5) %>%
  summarise(Cytotoxic = mean(Cytotoxic1),Exhaution = mean(Exhaution1)) %>%
  # filter(RNA_snn_res.1.2 != 4) %>%
  ggscatter(x='Cytotoxic',y='Exhaution',
            label = 'cell_types_5',
            # color = "#00A6B3", shape = 21, size = 2,
            cor.coef = T,add = 'loess')+
  theme(panel.border = element_rect(fill=NA,color="black",linetype="solid",size = 1))
```

#### Fig3 d Vlnplot

```{r}
markers <- c('HAVCR2','LAG3','TIGIT','PDCD1','LAYN','TOX','CTLA4',
             'GNLY','IFNG','KLRD1','PRF1','GZMB','GZMH')
VlnPlot(se_ob_cd8.t,
        features = markers,pt.size = 0,stack = T,flip = T,
        group.by = 'condition')
```

#### Pseudotime analysis

```{r}
library(monocle3)

se_ob_cds <- subset(se_ob_t,(condition == 'ATC'&cell_types_6 =='CD8'))

cds <- SeuratWrappers::as.cell_data_set(se_ob_cds)
cds <- cluster_cells(cds = cds, reduction_method = "UMAP")
cds <- learn_graph(cds, use_partition = TRUE)
cds <- order_cells(cds,reduction_method = "UMAP")

plot_cells(
    cds = cds,
    color_cells_by = "pseudotime",
    show_trajectory_graph = F,group_label_size = 5,
)
cds_sub <- choose_graph_segments(cds)
```

```{r}
se_ob_t <- AddMetaData(
  object = se_ob_t,
  metadata = cds@principal_graph_aux@listData$UMAP$pseudotime,
  col.name = "pseudo.cd8.ATC"
)

```

### B cells

```{r}
source("~/Documents/mytools/seurat_help_func.R")
se_ob_imm <- seurat_read_rds('./data/PTC_ATC/se_ob_imm.rds',scale = 'part')
se_ob_b <- subset(se_ob_imm, cell_types_3 =='B')
```


#### Workflow

```{r}
# Workflow ----
se_ob_b <- Seurat::NormalizeData(se_ob_b, 
                               normalization.method = "LogNormalize", 
                               scale.factor = 10000) %>% 
  FindVariableFeatures(., 
                       selection.method = "vst", 
                       nfeatures = 2000)

# all.genes <- rownames(se_ob_b)
se_ob_b <- ScaleData(se_ob_b)

se_ob_b <- RunPCA(se_ob_b,
                   features = VariableFeatures(object = se_ob_b)[!str_detect(VariableFeatures(object = se_ob_b),'^IG.*')])

DimPlot(se_ob_b, reduction = "pca",group.by = 'orig.ident')
ElbowPlot(se_ob_b,ndims = 50)
VlnPlot(object = se_ob_b, features = "PC_1", group.by = "orig.ident", pt.size = .1)

library(harmony)
# Remove batch effect ----
se_ob_b <- se_ob_b %>% 
  RunHarmony("orig.ident", plot_convergence = TRUE)

library(cowplot)
p1 = DimPlot(se_ob_b, reduction = "harmony")
p2 = ElbowPlot(se_ob_b,reduction = "harmony",ndims = 30)
p3 = VlnPlot(object = se_ob_b, features = "harmony_1", group.by = "orig.ident", pt.size = .1)
plot_grid(p1,p2,p3)

se_ob_b <- RunUMAP(se_ob_b, 
                 reduction = 'harmony', 
                 dims = 1:20,
                 seed.use = 1) %>% 
  RunTSNE(.,reduction = 'harmony', 
          dims = 1:20,
          seed.use = 1) %>% 
  FindNeighbors(., reduction = 'harmony', 
                dims = 1:20)

se_ob_b <- FindClusters(se_ob_b, 
                      resolution = c(0.2, 0.4, 0.6,
                                     0.8, 1.0, 1.2)) 
```

#### Find markers

```{r}
b.markers <- FindAllMarkers(se_ob_b,
                            test.use = 'MAST',
                            only.pos = T)
```

#### Annotation Marker Gene

```{r}
b_marker = c('CRIP1','IGHM','MS4A1', #B 
             'SDC1', #plasma
             'IGHD', 'FCER2', 'TCL1A','IL4R',#naive
             'CD27', 'AIM2','TNFRSF13B', #memory
             # 'LRMP', 'SUGCT', 'MME', 'MKI67','AICDA',#GC
             'STMN1'#,'PCNA' #proliferating
             
)
plasma_marker = c('IGHG1','IGHG2','IGHG3', #IGG
                  'IGHA1','IGHA2','JCHAIN', #IGA
                  'MZB1','DERL3','TNFRSF17' #plasma
)

jjDotPlot(object = se_ob_b,
          gene = c(b_marker,plasma_marker),
          xtree = F,
          ytree = T, 
          # rescale = T, rescale.min = 0, rescale.max = 1,
          id = 'RNA_snn_res.1')

DimPlot(se_ob_b,
        group.by = 'RNA_snn_res.1',
        label = T,
        cols = randomcoloR::randomColor(24))

DimPlot(se_ob_b,
        group.by = 'RNA_snn_res.1.2',
        split.by = 'orig.ident',
        label = T,
        cols = randomcoloR::randomColor(24))
```

#### Annotation Cell Type

```{r}
naive <- c(6,12)
memo <- c(0,14)
prolifer <- c(2)
igg <- c(3,7,9,10,13,15,16,17,18)
iga <- c(4,1,8)
other <- c(5,11)

cell_types_6 <- c(rep('Naive',length(naive)),
                 rep('Memory',length(memo)),
                 rep('Prolifering',length(prolifer)),
                 rep('IgG',length(igg)),
                 rep('IgA',length(iga)),
                 rep('other',length(other)))

cell_type_5 <- c('Naive-C1','Naive-C2',
                 'Memory-C1','Memory-C2',
                 'Prolifering-C1',
                 'IgG-C1','IgG-C2','IgG-C3','IgG-C4','IgG-C5',
                 'IgG-C6','IgG-C7','IgG-C8','IgG-C9',
                 'IgA-C1','IgA-C2','IgA-C3',
                 "Others","Others")
cell_map <- data.frame(clusters = c(naive,memo,prolifer,igg,iga,other),
                       cell_types_6 = cell_types_6,
                       cell_types_5 = cell_type_5) %>% 
  arrange(clusters)


cluster_letters <- cell_map$cell_types_6[se_ob_b$RNA_snn_res.1.2]
se_ob_b <- AddMetaData(
  object = se_ob_b,
  metadata = cluster_letters,
  col.name = 'cell_types_6')

cluster_letters <- cell_map$cell_types_5[se_ob_b$RNA_snn_res.1.2]
se_ob_b <- AddMetaData(
  object = se_ob_b,
  metadata = cluster_letters,
  col.name = 'cell_types_5')
se_ob_b <- subset(se_ob_b,cell_types_6 != "other")
DimPlot(se_ob_b,
        group.by = 'cell_types_6',
        label = T,label.size = 3.5,
        cols = randomcoloR::randomColor(17))

jjDotPlot(object = se_ob_b,
          gene = c(b_marker,plasma_marker),
          xtree = F,
          ytree = F, 
          rescale = T, rescale.min = 0, rescale.max = 1,
          id = 'cell_types_6')
```

#### Cell Fraction

```{r}
cell_ident <- data.frame(orig.ident = se_ob_b$orig.ident,
                        sample = se_ob_b$condition,
                        cluster = se_ob_b$RNA_snn_res.1.2)

plot.data <- cell_ident %>% 
  group_by(orig.ident,sample,cluster) %>% 
  summarise(count = n()) %>% 
  ungroup() %>% 
  group_by(orig.ident) %>% 
  mutate(freq = round(count/(sum(count)),3)) %>% 
  ungroup() 

# bar plot
library(ggpubr)
ggboxplot(plot.data,'cluster','freq',
          color = 'sample',add = 'jitter')+
  rotate_x_text(40)+
  stat_compare_means(aes(group = sample))
```

### Macrophage
```{r}
source("~/Documents/mytools/seurat_help_func.R")
se_ob_imm <- seurat_read_rds('./data/PTC_ATC/se_ob_imm.rds',scale = 'part')
se_ob_my <- subset(se_ob_imm, cell_types_3 == 'Myeloid')
```
#### Workflow
````{r}
# Workflow ----
se_ob_my <- Seurat::NormalizeData(se_ob_my, 
                               normalization.method = "LogNormalize", 
                               scale.factor = 10000) %>% 
  FindVariableFeatures(., 
                       selection.method = "vst", 
                       nfeatures = 2000)

# all.genes <- rownames(se_ob_b)
se_ob_my <- ScaleData(se_ob_my)

se_ob_my <- RunPCA(se_ob_my,
                   features = VariableFeatures(object = se_ob_my))

DimPlot(se_ob_my, reduction = "pca",group.by = 'orig.ident')
ElbowPlot(se_ob_my,ndims = 50)
VlnPlot(object = se_ob_my, features = "PC_1", group.by = "orig.ident", pt.size = .1)

library(harmony)
# Remove batch effect ----
se_ob_my <- se_ob_my %>% 
  RunHarmony("orig.ident", plot_convergence = TRUE)

library(cowplot)
p1 = DimPlot(se_ob_my, reduction = "harmony")
p2 = ElbowPlot(se_ob_my,reduction = "harmony",ndims = 30)
p3 = VlnPlot(object = se_ob_my, features = "harmony_1", group.by = "orig.ident", pt.size = .1)
plot_grid(p1,p2,p3)

se_ob_my <- RunUMAP(se_ob_my, 
                 reduction = 'harmony', 
                 dims = 1:20,
                 seed.use = 1) %>% 
  RunTSNE(.,reduction = 'harmony', 
          dims = 1:20,
          seed.use = 1) %>% 
  FindNeighbors(., reduction = 'harmony', 
                dims = 1:20)

se_ob_my <- FindClusters(se_ob_my, 
                      resolution = c(0.2, 0.4, 0.6,
                                     0.8, 1.0, 1.2)) 
```

#### Find markers

```{r}
my.markers <- FindAllMarkers(se_ob_my,
                            test.use = 'MAST',
                            only.pos = T)
```

#### Annotation Marker Gene

```{r}
dc = c('LGALS2','LAMP3','CD1E','CD1C')
m1 = c('IL23A','IL1A','IL1B','TNF')
m2 = c('CD163','PDCD1LG2','MRC1','CCL13','CCL18','LYVE1','VEGFB','CTSB','MMP14','FN1')
mono = c('S100A12','VCAN','CXCL8','FCN1','CD300E','THBS1','G0S2','NAMPT','BCL2A1','CD36','MARCO',
         'AQP9','TIMP1','CCR2','LILRA5')
markers = c(dc,m1,m2,mono)
my_marker = c('CD1E','CD1C','LAMP3','S100A12','VCAN','TNF','IL23A','CD163','MRC1')

jjDotPlot(object = se_ob_my,
          gene = my_marker,
          ytree = T,
          rescale = T, rescale.min = 0, rescale.max = 1,
          id = 'RNA_snn_res.1.2')

AverageHeatmap(object = se_ob_my,
               markerGene = na.omit(markers),cluster_columns = T,
               group.by = 'RNA_snn_res.1.2')

DimPlot(se_ob_my,
        group.by = 'RNA_snn_res.1.2',
        label = T,
        cols = randomcoloR::randomColor(24))

DimPlot(se_ob_my,
        group.by = 'RNA_snn_res.1.2',
        split.by = 'orig.ident',
        label = T,
        cols = randomcoloR::randomColor(24))
```

#### Cal gene sig
```{r}
GeneSets = as.list(readxl::read_excel('~/Documents/bio-ann-file/myeloid_sig.xlsx'))

for (type in names(GeneSets)) {
  genes = GeneSets[[type]]
  genes = genes[!is.na(genes)]
  se_ob_my <- AddModuleScore(
    object = se_ob_my,
    features = list(genes),
    name = type)
}

jjDotPlot(object = se_ob_my,
          gene = c('DC.activation1',
                   "DC.migration1",
                   "DC.tolorgenic1",
                   'Monocyte.signature1',
                   'M1.signature1',
                   'M2.siganture1'),
          ytree = F,
          rescale = T, rescale.min = 0, rescale.max = 1,
          id = 'RNA_snn_res.1.2')+coord_flip()
```

#### Cell Fraction

```{r}
cell_ident <- data.frame(orig.ident = se_ob_my$orig.ident,
                        sample = se_ob_my$condition,
                        cluster = se_ob_my$cell_types_6)

plot.data <- cell_ident %>% 
  group_by(orig.ident,sample,cluster) %>% 
  summarise(count = n()) %>% 
  ungroup() %>% 
  group_by(orig.ident) %>% 
  mutate(freq = round(count/(sum(count)),3)) %>% 
  ungroup() 

# bar plot
library(ggpubr)
ggboxplot(plot.data,'cluster','freq',
          color = 'sample',add = 'jitter')+
  rotate_x_text(40)+
  stat_compare_means(aes(group = sample))
```


#### Annotation Cell Type

```{r}
dc <- c(6,13)
mono <- c(2,7,14)
m1 <- c(3,4,9,10)
m2 <- c(0,5,8,11,12)
t <- c(1)

cell_types_6 <- c(rep('DC',length(dc)),
                 rep('Mono',length(mono)),
                 rep('M1',length(m1)),
                 rep('M2',length(m2)),
                 rep('T',length(t)))

cell_type_5 <- c('DC-C1','DC-C2',
                 'Mono-C1','Mono-C2','Mono-C3',
                 'M1-C1','M1-C2','M1-C3','M1-C4',
                 'M2-C1','M2-C2','M2-C3','M2-C4','M2-C5',
                 "T")
cell_map <- data.frame(clusters = c(dc,mono,m1,m2,t),
                       cell_types_6 = cell_types_6,
                       cell_types_5 = cell_type_5) %>% 
  arrange(clusters)


cluster_letters <- cell_map$cell_types_6[se_ob_my$RNA_snn_res.1.2]
se_ob_my <- AddMetaData(
  object = se_ob_my,
  metadata = cluster_letters,
  col.name = 'cell_types_6')

cluster_letters <- cell_map$cell_types_5[se_ob_my$RNA_snn_res.1.2]
se_ob_my <- AddMetaData(
  object = se_ob_my,
  metadata = cluster_letters,
  col.name = 'cell_types_5')
se_ob_my <- subset(se_ob_my,cell_types_6 != "T")
DimPlot(se_ob_my,
        group.by = 'cell_types_6',
        label = T,label.size = 3.5,
        cols = randomcoloR::randomColor(4))
```

#### ssGSVA

```{r}

library(Seurat)
library(msigdbr)
library(GSVA)
library(clusterProfiler)
library(patchwork)

scRNA <- subset(se_ob_my,cell_types_6 == 'M2')
DefaultAssay(scRNA) <- "RNA"
scRNA <- NormalizeData(scRNA)

Idents(scRNA) <- "condition" 
expr <- AverageExpression(scRNA, assays = "RNA", slot = "data")[[1]]
expr <- expr[rowSums(expr)>0,]  #选取非零基因
expr <- as.matrix(expr)

## geneset1
genesets <- msigdbr::msigdbr(species = "Homo sapiens", category = "H") 
genesets <- subset(genesets, select = c("gs_name","gene_symbol")) %>% 
  as.data.frame() %>% 
  mutate(gs_name = str_remove(gs_name,'HALLMARK_'))
seleted_pathway = c('APOPTOSIS','HYPOXIA','EPITHELIAL_MESENCHYMAL_TRANSITION',
                    'GLYCOLYSIS','ANGIOGENESIS')
genesets <- genesets %>% dplyr::filter(gs_name %in% seleted_pathway)
genesets <- split(genesets$gene_symbol, genesets$gs_name)

## geneset2
genesets2 <- readxl::read_excel('~/Documents/bio-ann-file/2499_Immune_Related_Gene.xlsx') 
# genesets2$Symbol <- str_to_title(genesets2$Symbol)
genesets2 <- split(genesets2$Symbol,genesets2$Category)
# genesets2$BCRSignalingPathway = NULL
# genesets2$TCRsignalingPathway = NULL
# genesets2$NaturalKiller_Cell_Cytotoxicity = NULL

genesets = c(genesets,genesets2)

gsva.res <- GSVA::gsva(expr, genesets, method="gsva") 

p_macro = pheatmap::pheatmap(gsva.res, 
                             show_colnames = T,show_rownames = T,
                             scale = "none",
                             cluster_cols=T,
                             cluster_rows = T,
                             main = 'M2',
                             # labels_col = c('nThyroid','PTC','ATC')
                             )
```